A Computational Account of Default-Mode Function\\by Control Theory and Reinforcement Learning

### 2016-08-05 paper.tex
stuf

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{mdframed}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{wrapfig}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{multirow}

@@

vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 \#3750} individual. Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled. Thus, dorsal regions of rostral mPFC contain value representations in the frame of reference of a modeled (as opposed to an actuated) preference irrespective of whether this applies to another’s or to one’s own likely actions and goals. Likewise, ventral mPFC contains a representation of value in the frame of reference of an executed choice, even if this executed choice reflects one’s own or another’s preferences. dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand. Critically, we also show this activity. While the simplest interpretation of this effect is that the region is simulating one’s own, currently irrelevant, preferences an alternative possibility is that the activity is projec- ting one’s own values into the mind of the partner, while simu- lating the partner’s choices. In essence, estimating the extent to which my own values would influence my partner, if they were making the choice. This iterated reasoning would be.

@@

Simony 2016
This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system

@@

@@

\item
 empirical evidence hinting at a two-process function of the hippocampus comes from structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al. 2006b). Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al. 2006b; Woollett \& Maguire 2009).

@@

\section{Methods}

\subsection{Linear control network model for brain organization}
Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
\begin{equation}
  \dot{\x}(t) = \A\x(t) + \B\u(t).
  \label{eq:lti}
\end{equation}
Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node). The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}. A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
\begin{equation}
  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
\end{equation}
has full rank, i.e
\begin{equation}
  rank(\C) = n.
\end{equation}

@@

\subsection{Optimal control: an LQR feedback controller}
We propose to use a linear quadratic regulator (LQR) for the feedback controller. Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
\begin{equation}
  \begin{split}
    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
  \end{split}
  \label{eq:hj}
\end{equation}
where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively. A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
\begin{equation}
  \u(t) = -\K(t)\x(t),
\end{equation}
where
\begin{equation}
  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
\end{equation}
and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
\begin{equation}
\begin{split}
&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
\end{split}
\end{equation}
%% \subsection{Minimal control energy controller}
\subsubsection{Related works}
The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...

\subsection{Demystifying the free-energy principle and active-inference}
In this section we will develop from first-principles, the bare-bones minimalistic ideas needed to build a free-energy principle for general decision-making. This ideas were first developed by Hinton et al. in the early 90s in the form of the so-called  \textit{Helmholtz machine}. Theories like Friston's free-energy principle and active-inference will then emerge as particular instances of this general framework, with particular design choises. For example the Friston theory axiomatizes that the brain uses a (problematic, as it implicitly assumes that posterior of each hidden unit is factorial, e.g) wake-sleep algorithm to train the underlying Helmholtz machine, etc.
\begin{table}[H]
  \begin{tabular}{p{2cm}|p{11cm}}
         \hline
         \textbf{variable}    & \textbf{short description}  \\ \hline
         $\langle E\rangle_p$ & Expectation (a.k.a average) of the
         random quantity $E$ w.r.t to the probability density $p$.\\ \hline
         $\mathcal H$ & The usual information-theoretic entropy operator.\\ \hline
         $D_{KL}(p||q)$ & The Kullback-Leibler divergence between the probability densities $p$ and $q$. \\ \hline
             $\zeta$ & Observations. In Friston's free-energy principle this has a decomposition in to two terms: the brain's internal state $b$ and sensory inputs $s$, i.e $\zeta = (s, b).$ \\ \hline
             $\psi$ & Hidden variables. This should be understood as the unobservable states of the external environment (to which the brain is trying to adapt by learning).\\ \hline
             $p_G(.|\zeta)$ & Generative density for ...\\ \hline
         $p_R(.|\zeta)$ & Recognition density for ... Does some kind of predictive coding (?).\\ \hline
         $F_G(\zeta)$ & Helmholtz free-energy for a model $p_G$ of generating the observation $\zeta$. This measures the surprise incured upon observing $\zeta$ generated by the model $G$.\\ \hline
         $F^R_G(\zeta)$ & Variational Helmholtz free-energy from model $R$
         to $G$.  Note that $F^G_G = F_G$.\\ \hline
  \end{tabular}
  \caption{Table of notations.}
\end{table}
\subsubsection{Helmholtz free-energy and the generative model}
Our starting point will be to build an approximation $p_G$ for the true density $p$ of the observations, so that this approximate density corresponds to the partition function of thermodynamic system. So,
\begin{eqnarray}
  \begin{split}
    \text{generative surprise } &= -\log(p_G(\zeta)) = -\log(p_G(\zeta)) \times 1 = -\log(P_G(\zeta))\sum_{\psi}p_G(\psi |\zeta)\\
    &= -\sum_{\psi}p_G(\psi, \zeta)\log(p_G(\zeta))
    =-\sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi, \zeta)/p_G(\psi|\zeta))\\
    &= \sum_{\psi}p_G(\psi |\zeta)\log(p_G(\psi|\zeta))-\sum_{\psi}p_G(\psi |\zeta)\log  (p_G(\psi, \zeta))\\
    &= -\langle \log  (p_G(., \zeta)) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))\\
    &= \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta)) := F_G(\zeta)
  \end{split}
  \label{eq:helm}
\end{eqnarray}
where $E_G(\psi, \zeta)$ is the energy of an fictive thermodynamic system defined by setting
\begin{equation}
  E_G(\psi, \zeta) := -\log(p_G(\psi, \zeta)).
  \label{eq:gibbs}
\end{equation}
%% In the above, the variable $\psi$ is a hidden variable, and can / should be interpreted as the unobservable state of the external world.
The last quantity $F_G(\zeta) := \langle E_G(., \zeta) \rangle_{p_G(. |\zeta)} - \mathcal H(p_G(. |\zeta))$ in \eqref{eq:helm} is nothing other than the Helmholtz free-energy for the physical system correspond to the external world, prescribed by the generative density $p_G(\psi|\zeta)$ for the occupation of \textit{macrostates} $\psi$. Thus generative surprise and generative Helmholtz free-energy are different views on exactly the same object.

The goal of the brain is then to optimize over the generative model $G$ (to iteratively or analytically modify the generative density $p_G(.|s)$, so as to minimize their surprise. It turns out that a direct attempt to attack this optimization problem by gradient descent on the free-energy $F_G(s)$ is furtile: the parameter update steps are not ``very clean'', and require rather cumbersome and heavy computations. A workaround is then to introduce a second density $p_R(.|\zeta)$ called a \textit{recognition} density to work in tandem with the generative density $p_G(.|\zeta)$: the former dreams / fantacizes whilst the latter tries to generate sensations which match these dreams! This primal-dual idea, first proposed in Hinton et al. 1995, is at the heart ot the general free-energy principle that we'll introduce shortly.

\subsubsection{Variational Helmholtz free-energy and the recognition density}
In this subsection we'll develop a very insightful upper bound for the generative surprise (i.e generate Helmholtz free-energy), called the \textit{variational} (Helmholtz) free-energy. As an avant-gout of what is to come shorty, let's just that the well-known \textit{free-energy principle} is simply a workaround whereby the minimization surprise (intractable) is replaced with the  minimization a carefully chosen upper bound thereof.

Now, invoking \eqref{eq:gibbs} and applying Bayes rule, we get the Gibbs
distribution
\begin{equation}
  p_G(\psi|\zeta) = \frac{p_G(\psi|\zeta)}{p_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)} = \frac{\exp(-E_G(\psi, \zeta))}{Z_G(\zeta)},
\end{equation}
where $Z_G(\zeta) := \log(p_G(\zeta)) = \sum_{\psi'}\exp(-E_G(\psi',\zeta))$, the normalizing partition function for the model \ref{eq:gibbs}.
Whence $\forall \psi$, $Z_G(\zeta) = p_G(\zeta) = \exp(-E_G(\psi, \zeta)) / p_G(\psi|\zeta)$, and so we have the invariance relation
\begin{equation}
  -\log(Z_G(\zeta)) = -\log(p_G(\zeta)) = F_G(\zeta) = E_G(\psi, \zeta) + \log(p_G(\psi|\zeta)).
  \end{equation}
Now, in the equation above, the LHS only depends on the generative model $G$ and the data point $\zeta$: it doesn't depend on the hidden variable $\psi$, etc. So, taking expectations w.r.t an arbitrary density\footnote{The conditioning in $P_R(.|\zeta)$ is because this density is selected from a world in which the sensory inputs and internal brain state vector $\zeta$ is assumed already observed.} $p_R(.|\zeta)$ yields
\begin{equation}
  \begin{split}
    F_G(\zeta) &= -\log(Z_G(\zeta)) = \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} + \sum_{\psi}p_R(\psi|\zeta)\log(p_G(\psi|\zeta))\\
    &= \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta)) - \sum_{\psi}p_R(\psi|\zeta)\log(p_R(\psi|\zeta)/p_G(\psi|\zeta))\\
    &= F^R_G(\zeta) - D_{KL}(P_R(.|\zeta) || P_G(.|\zeta)),
  \end{split}
  \label{eq:fe}
\end{equation}
where $F^R_G(\zeta)$ is the \textit{variational} Helmholtz free-energy from $R$ to $G$ defined by
\begin{equation}
  F^R_G(\zeta) := \langle E_G(., \zeta)\rangle_{P_R(.|\zeta)} - \mathcal H(p_R(.|\zeta))
\end{equation}
and $D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))$ is the Kullback-Leibler divergence between the $p_R(.|\zeta)$ and the generative density $p_G(.|\zeta)$. Note that $F^G_G = F_G$.

\subsubsection{A general free-energy principle}
We can resume the situation as follows:
\begin{mdframed}
\begin{equation}
  \begin{split}
    \text{generative surprise } &:= -\log(p_G(\zeta)) = \underbrace{F_G(\zeta)}_{\text{generative free-energy}} \\
    &=\underbrace{F^R_G(\zeta)}_{\text{variational free-energy}} - \underbrace{D_{KL}(P_R(.|\zeta) || P_G(.|\zeta))}_{\text{complexity}} \\
    &\le F^R_G(\zeta),
    \text{ with equalitity if }p_R(.|\zeta) = p_G(.|\zeta),
    \end{split}
\end{equation}
\end{mdframed}
where we have used the fact that KL divergence is always nonnegative. In fact, we have the following theorem
\begin{theorem}
  It holds that
  $$
  \text{Minimal generative surprise = minimal (Helmholtz) variational free-energy},$$
  i.e
  \begin{equation}
    \min_{G}F_G(\zeta) = \min_{G,R}F^R_G(\zeta)
    \end{equation}
\end{theorem}

\subsubsection{Helmholtz machines}
% \begin{mdframed}
  \textbf{Assumption:} In both generative and recognition components of the network, there is conditional independence of  neurons in the same layer, given the data. Precisely
$$
p_G(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_G(\psi_k^{(l)} | \zeta),\hspace{1cm}
p_R(\psi^{(l)}|\zeta) = \Pi_{k=1}^{h_l}p_R(\psi_k^{(l)} |\zeta)
$$
% \end{mdframed}
\subsection{The wake-sleep algorithm}
...
\subsection{Friston's active-inference}
This is nothing but an application of the Dayan's wake-sleep algorithm for training a Helmholtz machine model of the brain...

The following critics can be made:
\begin{itemize}
  \item 
    As noted by Dayan et al. (\textit{Variants of Helmholtz machines}), the inter-neuronal intra-layer independence which is at the center of the HM becomes severely problematic as it is agnostic to the known organization of cortical layers...
    \end{itemize}

\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
\begin{itemize}
  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
    \end{itemize}
The material presented is a revisit of \cite{braun2011path}. See also \cite{ortega2013thermodynamics}
\paragraph{Utility functions and conjugate pairs.}
  Let $(\Omega, \mathcal F, P)$ be a probablity space. A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
  \begin{itemize}
  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
    $A,B \in \mathcal F$.
  \item monotonicity: $f_{U}$ is strictly increasing.
    \end{itemize}

@@

  \paragraph{Free-utility functional.}
    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$. Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
    \begin{equation}
      J(P'|U, P) = \langle U \rangle_{P'} + \alpha \mathcal H(P'),
    \label{eq:free_u}
    \end{equation}
    where
    \begin{equation}
      \mathcal H(P') := \langle \log(P') \rangle_{P'} := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))
    \end{equation}
    is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
  \begin{equation}
    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
  \end{equation}
  In particular, if $P$ is the Gibbs measure at temperature $\alpha$ corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
  \begin{equation}
    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
    \end{equation}

  \paragraph{The free-energy / utility principle (of Friston ?).}
  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$. We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.  The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$. The difference in free-utility is then
  \begin{equation}
    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\langle \Delta U\rangle_{P_{\text{fin}}}}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
    \label{eq:free}
  \end{equation}
  where
  \begin{equation}
    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \langle \log(\P_{\text{fin}}/\P_{\text{ini}})\rangle_{\P_{\text{fin}}} := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
  \end{equation}
  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
a trade-off between an expected utility $\Delta U$ (first term) and
the information cost of transforming Pi
into $P_{\text{fin}}$ (second
term). In this interpretation $P_{\text{ini}}$ represents an initial choice
probability or policy, which includes the special case of the
uniform distribution where the decision-maker has initially no
preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
considers the utility constraint U∗
that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
\begin{equation}
P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
  \end{equation}

@@

Two limit cases are worth considering.
\begin{itemize}
\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \langle \Delta U\rangle_{P'_{\text{fin}}}$, and so it's optimal to take
  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
\end{itemize}

To conclude this section, let's note that \cite{braun2011path} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework. For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...

@@

\section{Relation to Other Statistical Models (Elvis)}
\paragraph{Control Theory.}
abc

\paragraph{Reinforcement Learning.}
abc

\paragraph{Free Energy Principle.}
The free-energy principle in it's present form (including notions like ``generative density'', ``recognition density'', etc.) can be traceback to works of Dayan \& Hinton \cite{dayan1995helmholtz} in which the introduced the so-called Helmholtz machine...

\begin{itemize}
  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement-learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
  \end{itemize}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{wrapfig}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{ulem}
\usepackage{multirow}

@@

vmPFC regions respond more to the value difference that will be acted on in the current trial (executed in red), and dmPFC to the currently irrelevant value difference (modeled in blue) {Nicolle, 2012 #3750} individual. Instead activity in vmPFC reflects a choice preference that is executed and rostral dmPFC a choice preference that is modeled. Thus, dorsal regions of rostral mPFC contain value representations in the frame of reference of a modeled (as opposed to an actuated) preference irrespective of whether this applies to another’s or to one’s own likely actions and goals. Likewise, ventral mPFC contains a representation of value in the frame of reference of an executed choice, even if this executed choice reflects one’s own or another’s preferences. dmPFC that reflects the values and preferences of another individual (here temporally discounted at a rate specific to the individual), even when they are not directly relevant to the task at hand. Critically, we also show this activity. While the simplest interpretation of this effect is that the region is simulating one’s own, currently irrelevant, preferences an alternative possibility is that the activity is projec- ting one’s own values into the mind of the partner, while simu- lating the partner’s choices. In essence, estimating the extent to which my own values would influence my partner, if they were making the choice. This iterated reasoning would be.

@@

Simony 2016
This implies that the processing timescale of the DMN is longer than the length of single paragraphs (B40s) and that information accumulated from prior paragraphs can affect ongoing processing in this system
Our results suggest that the DMN accumulates and integrates information over minutes. Stimulus-induced correlation patterns in the DMN were largest for the intact story, greatly decreased in the paragraph scramble condition, and were almost entirely abolished in the word scramble condition.

@@

\item
Kai had an interesting idea: the deja-vu effect could be when you already accidentally perturbed past such that it really occures in reality later

@@

\item
 empirical evidence hinting at a two-process function of the hippocampus comes from structural MRI studies of expert navigators (London taxi drivers) who show increased grey matter volume in posterior hippocampus seemingly at the expense of reduced grey matter volume in anterior hippocampus (Maguire et al. 2006b). Moreover, their increased spatial knowledge appears to come at a cost to the acquisition of new visual associative information (Maguire et al. 2006b; Woollett & Maguire 2009).

@@

\section{Methods}
\subsection{Linear control network model for brain organization}
Consider the following linear time-invariant (LTI) dynamical system as a (toy) model for high-level brain function
\begin{equation}
  \dot{\x}(t) = \A\x(t) + \B\u(t).
  \label{eq:lti}
\end{equation}
Here, the $n$-by-$n$ matrix $\A$ denotes a model of the brain's wiring (for example a resting state connectome computed from an anatomical atlas), while the $n$-by-$k$ ``input matrix'' $\B$ describes which nodes can be controlled by us, an external \textit{controller}, via medical intervention or a careful choice of stimulus presentation, for example.
At time $t \ge 0$, let  the $n$-vector  $\x(t) := (\x_1(t),\ldots,\x_n(t))$ encodes the state of the network (one value for each node). The aim is to supply values of $\u_1(t),\ldots,\u_k(t)$ for $k \le n$ controls as a function of the time $t$,  to take the system from any prescribed initial state $\x(0) = \x^{\text{init}} \in \mathbb R^n$ to any prescribed final state $\x(\tau) = \x^{\text{fin}} \in \mathbb R^n$ in finite time $\tau < \infty$.
When such a controlling is possible, we say that the system $(\A,\B)$ is \textit{controllable}. A precise sufficient and necessary condition for such controllability is the Kalman condition: $(\A,\B)$ is controllable iff the $n \times nk$ \textit{controllability matrix}
\begin{equation}
  \C := (\B|\A\B|\ldots|\A^{n-1}\B)
\end{equation}
has full rank, i.e
\begin{equation}
  rank(\C) = n.
\end{equation}

@@

\subsection{Optimal control: an LQR feedback controller}
We propose to use a linear quadratic regulator (LQR) for the feedback controller. Thus, consider the time-varying \text{value function} $V: [0, \tau] \times \mathbb R^n \rightarrow \mathbb R$, defined by the following Hamilton-Jacobi cost-to-go
\begin{equation}
  \begin{split}
    &V(t, \z) := \min_{\u} \frac{1}{2}\x_{\text{fin}}^T\Q_{\text{fin}}\x_{\text{fin}} + \int_{0}^\tau\left(\frac{1}{2}\x(t)^T\Q\x(t) + \frac{1}{2}\u(t)^T\R\u(t)\right)dt,\\
    &\text{subject to } \x(t) = \z, \dot{\x}(t) = \A\x(t) + \B\u(t),
  \end{split}
  \label{eq:hj}
\end{equation}
where the matrices $\Q$ and $\R$, precised by design considerations and subject to meta-optimization, are restricted to be positive semi-definite and positive definite, respectively. A classical calculation (reminiscent of the \textit{Pontryagin minimization principle}) reveals that the optimal feedback control in \eqref{eq:hj} is given by
\begin{equation}
  \u(t) = -\K(t)\x(t),
\end{equation}
where
\begin{equation}
  \K(t) := \R^{-1}\B^T\P(t)\text{ (Kalman gain matrix)}
\end{equation}
and the time-varying positive semi-definite matrix $\P(t)$ solves following differential Riccati equation (DRE)
\begin{equation}
\begin{split}
&-\dot{\P}(t) = \A^T\P(t) + \P(t)\A - \P(t)\B\R^{-1}\B^T\P(t) + \Q\\
&\text{subject to }\P(\tau) = \Q_{\text{fin}}.
\end{split}
\end{equation}
\subsection{Minimal control energy controller}
\subsection{Related works}
The model proposed in \cite{betzel2016} can be cast in the form \eqref{eq:hj}, with the particular choice $\Q = \textbf{I} = \rho^{-1}\R$ and $\Q_{\text{fin}} = \textbf{0}$, but with a rather ad-hoc handle on stability issues and choice driver nodes...

\subsection{A thermodynamic model for bounded rationality, aka robust optimality}
\begin{itemize}
  \item Recall that an agent is said to have \textit{bounded rationality} if they must take into account the cost of finding solutions to problems, and not just the utility of the final state.
  \item For example, consider an agent that must operate under a limited lifetime and/or computation cost.
  \item This is in contrast to agents with \textit{unbounded rationality} considered in classical game theory.
    \end{itemize}
The material presented is a revisit of \textcolor{blue}{[Braun 2011]}.
\paragraph{Utility functions and conjugate pairs.}
  Let $(\Omega, \mathcal F, P)$ be a probablity space. A function $U: \mathcal F \rightarrow \mathbb R$ is said to be a \textit{utility function} for this space if the conditional utility $U(A|B) := U(A \cap B) - U(B)$ has the following propertites:
  \begin{itemize}
  \item additivity: $U(A_1 \cap A_2 | B) = U(A_1|B) + U(A_2|B)$, for all events $A_1, A_2, B \in \mathcal F$.    
  \item statistic: there exists a function $f_{U} :\mathbb R_+ \rightarrow \mathbb R$ such that $U(A|B) = f(P(A|B))$, for all events
    $A,B \in \mathcal F$.
  \item monotonicity: $f_{U}$ is strictly increasing.
    \end{itemize}

@@

  \paragraph{Free-utility functional.}
    Let $(U, P)$ be a conjugate pair at temperature $\alpha > 0$  on a measurable space $(\Omega, \mathcal F)$. Given another probability measure $P'$ on the same space, define it's \textit{free utility} as
    \begin{equation}
      J(P'|U, P) = \mathbb E_{P'}(U) + \alpha \mathbb H(P'),
    \label{eq:free_u}
    \end{equation}
    where $\mathbb H(P') := -\sum_{\omega \in \Omega}P'(\omega)\log(P'(\omega))$ is the \textit{entropy} of $P'$ (measured in the Naperian base $e \approx 2.73$). It's not difficult to establish the upper bound
  \begin{equation}
    J(P'|U) \le J(P|U) = \sum_{\omega \in \Omega}U(\omega) =: U(\Omega).
  \end{equation}
  In particular, if $P$ is the Gibbs measure corresponding to $U$, then the upper bound above reduces to the \textit{log-partition function}
  \begin{equation}
    J(P'|U) \le U(\Omega) = -\alpha \log(Z_{U}(\alpha)).
    \end{equation}

  \paragraph{The free-energy / utility principle (of Friston ?).}
  We are now in shape to introduce the notion of free-energy for model transitions, and a variational principle for optimizing it. Consider thus an initial system described by a conjugate pair $(U_{\text{ini}}, P_{\text{ini}})$ at temperature $\alpha > 0$. We want to transform this to a new model by adding constraints represented by the utility function $\Delta U$.  The resulting system has final utility $U_{\text{fin}} = U_{\text{ini}} + \Delta U$. The difference in free-utility is then
  \begin{equation}
    \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} := J_{\text{fin}} - J_{\text{ini}} = \underbrace{\mathbb E_{P_{\text{fin}}}(\Delta U)}_{\textbf{accuracy}} -  \underbrace{\alpha D_{\text{KL}}(P_{\text{ini}}\|P_{\text{fin}})}_{\textbf{complexity}},
    \label{eq:free}
  \end{equation}
  where
  \begin{equation}
    D_{\text{KL}}(P_{\text{fin}}\|P_{\text{ini}}) := \sum_{\omega \in \Omega}\P_{\text{fin}}(\omega)\log(\P_{\text{fin}}(\omega)/\P_{\text{ini}}(\omega))
  \end{equation}
  is the Kullback-Leibler divergence, and represents the information cost (measured in energy units) of changing the initial system.
In the above formula, we've extensively used the fact that $(U_{\text{ini}}, P_{\text{ini}})$ is a congugate pair and so $U_{\text{ini}}(\omega) \equiv \alpha \log(P_{\text{ini}}(\omega))$ by virtue of \eqref{eq:boltzmann}.
  The two terms in \eqref{eq:free} (accuracy or expected gain in utility, and the complexity of the transition) can be viewed as dertiminants of bounded rational decision-making. They formalize
a trade-off between an expected utility $\Delta U$ (first term) and
the information cost of transforming Pi
into $P_{\text{fin}}$ (second
term). In this interpretation $P_{\text{ini}}$ represents an initial choice
probability or policy, which includes the special case of the
uniform distribution where the decision-maker has initially no
preferences between the different choices. The probability measure $P_{\text{fin}}$ is the final choice probability that we are looking for since it
considers the utility constraint U∗
that we want to optimize. We can then formulate a variational principle for bounded rationality in the probabilities $P_{\text{fin}}(\omega)$
\begin{equation}
P^*_{\text{fin}} := \underset{P_{\text{fin}}}{\text{argmax }} \Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}}
  \end{equation}

@@

Two limit cases are worth considering.
\begin{itemize}
\item \textbf{Low-temperature regime $\alpha \approx 0$:} Here $\Delta J_{(U_{\text{ini}}, P_{\text{ini}}) \rightarrow P_{\text{fin}}} \approx \mathbb E_{P'_{\text{fin}}}(\Delta U)$, and so it's optimal to take
  $$P^*_{\text{fin}}(\omega) \equiv \text{dirac}(\omega - \omega^*) = \begin{cases}1, &\mbox{if }\omega = \omega^*,\\0, &\mbox{ otherwise,}\end{cases}$$
  where $\omega^* := \argmax_{\omega \in \Omega}U(\omega)$. This corresponds to unbounded rational decision-making, in which the cost of transition / problem-solving is completely disregarded.
  \item \textbf{High-temperature regime $\alpha \rightarrow +\infty$:} In this limiting case, it's optimal to take $P^*_{\text{fin}}(\omega) \equiv P_{\text{ini}}(\omega)$, i.e the change is so costly that it's optimal to maintain the current choice probabilities.
\end{itemize}

To conclude this section, let's note that \textcolor{blue}{[Braun 2011]} show how their free-energy framework (on paths) links with the well-known Hamilton-Jacobi-Bellman optimal control framework. For example, one can re-derive the Linear Quadratic Gaussian (LQG) controller, which is a generalization of the LQR in \eqref{eq:hj}...

@@

\section{Relation to Other Statistical Models (Elvis)}
\paragraph{Control Theory}
abc

\paragraph{Q-Learning}
abc

\paragraph{Free Energy Principle}
abc
